Don't Mind the Gap: Bridging Network-wide Objectives and Device-level Configurations

### 2016-06-10 paper/v2/paper-edited.tex
first pass edits

We have implemented our language and compiler in a system called \sysname. To evaluate it, we use it to specify real policies of data center and backbone networks.
% and compared these specifications with existing configurations.
We find that our language expresses such policies easily, and that the compiler scales to topologies with hundreds of routers, compiling in under 9 minutes in all cases.

@@

To associate ranked paths with
one or more prefixes, we define a \sysname \emph{policy}.
Within a policy, statements with the form $t$\Path$p$
associate the prefixes defined by the predicate $t$ with the set of
ranked paths defined by the path expression $p$.  In general,
prefix predicates can be defined by arbitrary boolean combinations
(and, or, not) of concrete prefixes and community tags.  Here,
we assume we have already defined predicates \CD{PCust} for the
customer prefixes.
%and \CD{PInt} for our internal prefixes.
In the following code, ranked paths are associated with
customer prefixes,
and all other prefixes (\CD{true}). Policy statements are processed in
order with earlier policy statements taking precedence over later
policy statements. Hence, when the predicate \True{} follows
the statement involving \CD{PCust}, it is interpreted as
\CD{\True{} \AND{} \NOT{}PCust}.

@@

\subsection{Failure Analysis}

To implement path preferences in routing, BGP uses local preferences on a per-device basis. However, the distributed nature of BGP makes setting preferences locally to achieve a network-route routing policy difficult. This task becomes even more challenging in the presence of failures since routers running BGP lack a global view of the network.

@@

The problem of deciding local preferences for each router is closely reflected in the structure of the PGIR. In particular, whenever a given router appears in multiple different contexts in the PGIR (i.e., as multiple PGIR nodes), then the compiler must decide which context to prefer.
%
In our compilation example from Figure~\ref{fig:example-compilation}, the topology node $C$ can receive an advertisement from $E$ in context $(C,-,2)$ or it can receive an advertisement from its neighbor $D$ in context $(C,3,2)$. 
%
The compiler's task is to determine a \emph{total ordering} of contexts for each router in the topology, which reflects the relative preference of advertisements received in each context. For example, if the compiler determines that, for router $C$, context $(C,3,2)$ is preferred to context $(C,-,2)$, written as $(C,3,2) <_{pref} (C,-,2)$, then this means that $C$ will prefer advertisements heard from $D$ (with tag $(2,2)$) over $E$ (with tag $(-,2)$). Section~\ref{sec:abgp} discusses our encoding of preferences and filters in more detail.

@@

Algorithm~\ref{alg:failures} checks whether one context can be preferred to another ($N_1 <_{pref} N_2$). The algorithm walks from nodes $N_1$ and $N_2$ and ensures that for every \textit{step} $N_2$ can take to some new topology location, $N_1$ can, at the very least, also take a step to an equivalent topology location. When there is no such equivalent step, the algorithm attempts to take into account where the advertisement must have already traversed. In particular, it checks to see if there is an equivalent dominator (a topology location the advertisement must have passed through earlier) and walks from this new node instead. Algorithm~\ref{alg:failures} terminates since the number of related states $(x,y)$ that can be explored is finite.

@@

The checks for failure safety described above overlook one critical point: A better (lower rank) path might not be available due to loops. To avoid this situation, the compiler checks that, any time some context $N_1$ appears ``above'' (i.e., can reach) $N_2$ in the PGIR, then $N_1$ must be strictly preferred to $N_2$. The idea is that anytime $N_2$ is not usable because the advertisement already went through $N_1$, it ultimately does not matter since $N$ has a regret-free preference for $N_1$ anyway.

@@

\para{Aggregation-Induced Black Holes}
Because the \sysname compiler has the complete user policy and topology, it is also able to efficiently check that aggregates do not black hole traffic for up to $k$ failures. The algorithm for aggregation failure safety is detailed in Section~\ref{sec:black-holes} of the appendix.

@@

@@

\begin{figure}[t!]
%% \begin{lstlisting}[frame=single, mathescape=true]
%% $\Router{A}$
%%   Match $\PEER=C, \COMM=(3,2)$
%%     Export $\COMM \leftarrow (4,2), \MED \leftarrow 80, \PEER \leftarrow W$
%% $\Router{B}$
%%   Match $\PEER = C, \COMM = (-,2)$
%%     Export $\COMM \leftarrow (-,3), \COMM \leftarrow \text{noexport},$
%%            $\MED \leftarrow 81, \PEER \leftarrow W$
%%   Match $\PEER = C, \COMM = (3,2)$
%%     Export $\COMM \leftarrow (-,3), \COMM \leftarrow \text{noexport},$
%%            $\MED \leftarrow 81, \PEER \leftarrow W$
%% $\Router{C}$
%%   Match$[LP=99]$ $\PEER = E, \COMM = (-,2) $
%%     Export $\COMM \leftarrow (-,2), \PEER \leftarrow B$
%%   Match $\PEER = D, \COMM = (2,2)$
%%     Export $\COMM \leftarrow (3,2), \PEER \leftarrow A,B$
%% $\Router{D}$
%%   Match $\REGEX{X + Y}$
%%     Export $\COMM \leftarrow (2,2), \PEER \leftarrow C$
%% $\Router{E}$
%%   Match $\REGEX{Z}$
%%     Export $\COMM \leftarrow (-,2), \PEER \leftarrow C$
%% \end{lstlisting}
\begin{code}
  \Router{A}
    Match \PEER=C, \COMM=(3,2)
      Export \COMM \Arrow (4,2), \MED \Arrow 80, \PEER \Arrow W
  \Router{B}
    Match \PEER = C, \COMM = (-,2)
      Export \COMM \Arrow (-,3), \COMM \Arrow \text{noexport},
             \MED \Arrow 81, \PEER \Arrow W
    Match \PEER = C, \COMM = (3,2)
      Export \COMM \Arrow (-,3), \COMM \Arrow \text{noexport},
             \MED \Arrow 81, \PEER \Arrow W
  \Router{C}
    Match[LP=99] \PEER = E, \COMM = (-,2) 
      Export \COMM \Arrow (-,2), \PEER \Arrow B
    Match \PEER = D, \COMM = (2,2)
      Export \COMM \Arrow (3,2), \PEER \Arrow A,B
  \Router{D}
    Match \REGEX{X + Y}
      Export \COMM \Arrow (2,2), \PEER \Arrow C
  \Router{E}
    Match \REGEX{Z}
      Export \COMM \Arrow (-,2), \PEER \Arrow C
  \end{code}
  \vspace{-2em} 
  \caption{Abstract BGP router configurations. \label{fig:abgp-config}}
  \vspace{-1em}
\end{figure}

@@

We have written a prototype compiler for \sysname that is implemented in roughly 6700 lines of F\# code. The compiler includes command-line flags for enabling or disabling the use of the BGP MED attribute, AS path prepending, anycast, the no-export community, as well as for ensuring at least k-failure safety for user-specified aggregates. Since each prefix has a separate routing policy, the \sysname compiler is able to compile each routing policy in parallel. Currently \sysname supports generating Quagga router configurations out of the box. Users can also easily add new vendor-specific adaptors to translate from ABGP to other router configurations, or to incorporate the compiler into an existing template-based system, for example by mixing the \sysname configuration with other configuration components.

@@

All experiments are run on an 8 core, 3.6 GHz Intel Xeon processor running Windows 7.
%
Figure~\ref{fig:compilation-times} shows the compilation times for data centers (a), and backbone networks (b) of different sizes. For both policies, we measure the mean compilation time per prefix predicate since the compiler operates on each prefix predicate in parallel. Note that a single predicate can describe many prefixes, for example by matching on a disjunction of prefixes. At their largest sizes, the per-predicate compilation time is roughly 10 seconds for the data center network and 45 seconds for the backbone network.

@@

$ii)$ operators prevented unneeded propagation of more-specific route announcements from a neighboring AS based on their out-of-band knowledge about the topology, whereas \sysname propagated these advertisements; and

$iii)$ operators used a layer of indirection for community values, using community groups and re-writing values, to implement certain policies in a more maintainable manner, where \sysname uses flat communities.

@@

\section*{Appendix}
\label{sec:appendix}

\renewcommand\thesubsection{\Alph{subsection}}

\subsection{Aggregation-Induced Black Holes}
\label{sec:black-holes}

\begin{figure}[t!]
  \centering
  \includegraphics[width=\columnwidth]{figures/aggregation}
  \caption{Aggregation safety.}
  \label{fig:aggregation-safety}
  \vspace{-1em}
\end{figure}

As demonstrated with Example 2 in \S\ref{sec:motivation}, BGP aggregation can lead to subtle black-holing of traffic when failures occur. Deciding when this can happen requires knowledge of the routing policy and not just the topology. For instance, a policy might require all traffic for a particular prefix to go over a single link before being aggregated, even if there are several available in the toplogy. If that link fails, a black hole might be introduced. Fortunately, the PGIR captures the information about both the policy and topology precisely.
%, describing the how information meeting the policy flows over the topology.

We frame the aggregation problem as a problem of connectivity in the PGIR. Specifically, for each prefix that falls under an aggregate, we find a lower bound on the number of failures that would disconnect where the prefix originates from where its more specific aggregate is located. The difficulty is that the same links in the topology can appear in multiple places in the PGIR. We adopt the following simple strategy to lower bound the number of failures: $i)$ Pick a random start-to-end path in PGIR, $ii)$ remove all edges in the PGIR for the set of topological edges chosen, and $iii)$ repeat until no such path exists.
The number of PGIR paths that we are able to remove is a lower bound on the number of failures required to disconnect the prefix from its aggregate.

For example, imagine we use the data center example from \S\ref{sec:motivation}, with the policy: 
\CD{PG1} \Path \text{ }\End(\CD{A}), where \CD{PG1} falls under the \CD{PG} aggregate. Figure~\ref{fig:aggregation-safety} shows the network topology and a simplified PGIR. Since the compiler knows an aggregate will be placed at $X$, and it knows that, the route for \CD{PG1} will originate at $A$, we can compute the number of failures it would take to disconnect $A$ from $X$. We could remove the $A$--$D$--$X$ path first. We would then need to remove any other $A$--$D$ or $D$--$X$ links from the PGIR, though in this case there are none. Next, we could remove the links along the $A$--$C$--$X$ path, repeating the process. Because $A$ is then disconnected from $X$, the compiler knowns that 2 is a lower bound on the number of failures for aggregation safety for prefix \CD{PG1}. This process is repeated for other aggregation locations (e.g., $Y$).

>>>>

We have implemented our language and compiler in a system called \sysname. To evaluate it, we use it to specify real policies of data center and backbone networks.
% and compared these specifications with existing configurations.
We find that our language expresses such policies easily, and that the compiler scales to topologies with hundreds to thousands of routers, compiling in under 9 minutes in all cases.

@@

To associate ranked paths with
one or more prefixes, we define a \sysname \emph{policy}.
Within a policy, statements with the form $t$\Path$p$
associate the prefixes defined by the predicate $t$ with the set of
ranked paths defined by the path expression $p$.  In general,
prefix predicates can be defined by arbitrary boolean combinations
(and, or, not) of concrete prefixes and community tags.  Here,
we assume we have already defined the predicate \CD{PCust} for the
customer prefixes.
%and \CD{PInt} for our internal prefixes.
In the following code, ranked paths are associated with
customer prefixes,
and all other prefixes (\CD{true}). Policy statements are processed in
order with earlier policy statements taking precedence over later
policy statements. Hence, when the predicate \True{} follows
the statement involving \CD{PCust}, it is interpreted as
\CD{\True{} \AND{} \NOT{}PCust}.

@@

\subsection{Failure Safety}

To implement path preferences in routing, BGP uses local preferences on a per-device basis. However, the distributed nature of BGP makes setting preferences locally to achieve a network-wide routing policy difficult. This task becomes even more challenging in the presence of failures since routers running BGP lack a global view of the network.

@@

The problem of determining local preferences for each router is closely reflected in the structure of the PGIR. In particular, whenever a given router appears in multiple different contexts in the PGIR (i.e., as multiple PGIR nodes), then the compiler must decide which context to prefer.
%
In our compilation example from Figure~\ref{fig:example-compilation}, the topology node $C$ can receive an advertisement from $E$ in context $(C,-,2)$ or it can receive an advertisement from its neighbor $D$ in context $(C,3,2)$. 
%
The compiler's task is to determine a \emph{total ordering} of contexts for each router in the topology, which reflects the relative preference of advertisements received in each context. For example, if the compiler determines that, for router $C$, context $(C,3,2)$ is preferred to context $(C,-,2)$, written as $(C,3,2) <_{pref} (C,-,2)$, then this means that $C$ will prefer advertisements heard from $D$ (with tag $(2,2)$) over $E$ (with tag $(-,2)$). Section~\ref{sec:abgp} discusses our encoding of preferences and filters in more detail.

@@

Algorithm~\ref{alg:failures} checks whether one context can be preferred to another ($N_1 <_{pref} N_2$). The algorithm walks from nodes $N_1$ and $N_2$ and ensures that for every \textit{step} $N_2$ can take to some new topology location, $N_1$ can, at the very least, also take a step to an equivalent topology location. When there is no such equivalent step, the algorithm attempts to take into account where the advertisement must have already traversed. In particular, it checks to see if there is an equivalent dominator (a topology location the advertisement must have passed through earlier) and walks from this new node instead. 
At each step, it requires that current node reachable from $N_1$ has a path rank that is at least as good as that of the current node reachable from $N_2$ ($n_1 \leq_{rank} n_2$). The intuition here is that if $n_1 \nleq_{rank} n_2$, then we can likely fail every edge in the topology except for the path that leads to the current $n_1$ and $n_2$, thereby generating in a counter example.
Algorithm~\ref{alg:failures} terminates since the number of related states $(x,y)$ that can be explored is finite.

@@

The checks for failure safety described above overlook one critical point: A better (lower rank) path might not be available due to loops that are rejected by BGP. To avoid this situation, the compiler checks that, any time some context $N_1$ appears ``above'' (i.e., can reach) $N_2$ in the PGIR, then $N_1$ must be strictly preferred to $N_2$. The idea is that anytime $N_2$ is not usable because the advertisement already went through $N_1$, it ultimately does not matter since $N$ has a regret-free preference for $N_1$ anyway.

@@

\subsection{More Safety}

\para{Aggregation Safety}
\label{sec:agg-safety}

\begin{figure}[t!]
  \centering
  \includegraphics[width=\columnwidth]{figures/aggregation}
  \caption{Aggregation safety for a data center.}
  \label{fig:aggregation-safety}
  \vspace{-1em}
\end{figure}

As demonstrated with Example 2 in \S\ref{sec:motivation}, BGP aggregation can lead to subtle black-holing of traffic when failures occur. Deciding when this can happen requires knowledge of the routing policy and not just the topology. For instance, a policy might require all traffic for a particular prefix to go over a single link before being aggregated, even if there are several available in the toplogy. If that link fails, a black hole might be introduced. 
%
Because the \sysname compiler has the complete user policy and topology encoded in the PGIR, it is able to efficiently check that aggregates do not black hole traffic for up to $k$ failures.

We frame the aggregation problem as a problem of connectivity in the PGIR. Specifically, for each prefix that falls under an aggregate, we find a lower bound on the number of failures required to  disconnect where the prefix originates from where it is aggregated. The difficulty is that the same links in the topology can appear in multiple places in the PGIR. 
%
%
We adopt the following simple strategy to lower bound the number of failures: $i)$ Pick a random start-to-end path in PGIR, $ii)$ remove all edges in the PGIR for the set of topological edges chosen, and $iii)$ repeat until no such path exists.
The number of PGIR paths that we are able to remove is a lower bound on the number of failures required to disconnect the prefix from its aggregate.

For example, recall the data center example from \S\ref{sec:motivation}, with the policy: 
\CD{PG1} \Path \text{ }\End(\CD{A}), where \CD{PG1} falls under the \CD{PG} aggregate. Figure~\ref{fig:aggregation-safety} shows the network topology simplified PGIR. Since the compiler knows aggregation will occur at $X$, and it knows that the \CD{PG1} prefix will originate at $A$, we can compute the number of failures it would take to disconnect $A$ from $X$. We could remove the $A$--$D$--$X$ path first. We would then need to remove any other $A$--$D$ or $D$--$X$ links from the PGIR (in this case none). Next, we could remove the links along the $A$--$C$--$X$ path, repeating the process. Because $A$ is then disconnected from $X$, the compiler knowns that 2 is a lower bound on the number of failures for aggregation safety for prefix \CD{PG1}. This process is repeated for other aggregation locations (e.g., $Y$).

\para{Policy Safety}

Even when programming the network centrally, it is possible for operators to make mistakes and write incorrect policies.
\sysname includes many additional safety analysis checks to try to catch common mistakes at compile time. These include: (i) preference analysis to determine when backup paths/preferences will never be used, (ii) reachability analysis to check if topology locations that should be reachable according to the operator policy are not reachable after combining both the topology and policy, (iii) anycast analysis to find instances where the operator accidentally anycasts a prefix, (iv) aggregate analysis to find when aggregates do not summarize any specific prefix, and others.

@@

\begin{figure}[t!]
\begin{code}
  \Router{A}
    Match \PEER=C, \COMM=(3,2)
      Export \COMM \Arrow (4,2), \MED \Arrow 80, \PEER \Arrow W
  \Router{B}
    Match \PEER = C, \COMM = (-,2)
      Export \COMM \Arrow (-,3), \COMM \Arrow \text{noexport},
             \MED \Arrow 81, \PEER \Arrow W
    Match \PEER = C, \COMM = (3,2)
      Export \COMM \Arrow (-,3), \COMM \Arrow \text{noexport},
             \MED \Arrow 81, \PEER \Arrow W
  \Router{C}
    Match[LP=99] \PEER = E, \COMM = (-,2) 
      Export \COMM \Arrow (-,2), \PEER \Arrow B
    Match \PEER = D, \COMM = (2,2)
      Export \COMM \Arrow (3,2), \PEER \Arrow A,B
  \Router{D}
    Match \REGEX{X + Y}
      Export \COMM \Arrow (2,2), \PEER \Arrow C
  \Router{E}
    Match \REGEX{Z}
      Export \COMM \Arrow (-,2), \PEER \Arrow C
  \end{code}
  \vspace{-2em} 
  \caption{Abstract BGP router configurations. \label{fig:abgp-config}}
  \vspace{-1em}
\end{figure}

@@

@@

We have written a compiler for \sysname that is implemented in roughly 6700 lines of F\# code. The compiler includes command-line flags for enabling or disabling the use of the BGP MED attribute, AS path prepending, anycast, the no-export community, as well as for ensuring at least k-failure safety for user-specified aggregates. Since each prefix has a separate routing policy, the \sysname compiler is able to compile each routing policy in parallel. Currently \sysname supports generating Quagga router configurations out of the box. Users can also easily add new vendor-specific adaptors to translate from ABGP to other router configurations, or to incorporate the compiler into an existing template-based system, for example by mixing the \sysname configuration with other configuration components.

@@

All experiments are run on an 8 core, 3.6 GHz Intel Xeon processor running Windows 7.
%
Figure~\ref{fig:compilation-times} shows the compilation times for data centers (a), and backbone networks (b) of different sizes. For both policies, we measure the mean compilation time per prefix predicate since the compiler operates on each predicate in parallel. Note that a single predicate can describe many prefixes, for example by matching on a disjunction of prefixes. At their largest sizes, the per-predicate compilation time is roughly 10 seconds for the data center network and 45 seconds for the backbone network.

@@

$ii)$ operators prevented unneeded propagation of more-specific route announcements from a neighboring AS based on their out-of-band knowledge about the topology, whereas \sysname propagated these advertisements;

$iii)$ operators used a layer of indirection for community values, using community groups and re-writing values, to implement certain policies in a more maintainable manner, where \sysname uses flat communities; and

$iv)$ operators used BGP regular expression filters to enforce certain invariants that are independent of any particular prefix, whereas \sysname enforced these invariants per prefix.

@@
